Flattering Advice: Avoiding Disappointment in Advice-Giving

### 2025-01-09 manuscript/manuscript.qmd
Conversion to Quarto file

---
title: "Flattering Advice: Avoiding Disappointment in Advice-Giving"
shorttitle        : "Flattering Advice"
author: 
  - name          : "Amanda Chen"
    affiliation   : "1"
    corresponding : yes    
    email         : "zchengj@connect.ust.hk"
    address       : "Department of Management, The Hong Kong University of Science and Technology, Hong Kong"
  - name          : "David Hagmann"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "The Hong Kong University of Science and Technology"
note : |
  `r format(Sys.time(), "%e %B, %Y")`
abstract: |
  Good advice improves decision quality but often requires delivering unpleasant truths that may disappoint advisees. Across three pre-registered and incentivized experiments involving real adviser-advisee interactions (N = 3,900), we show that advisers prioritize avoiding disappointment at the expense of accuracy and their own earnings. In Study 1, advisers financially rewarded for accuracy still tailor recommendations to aspirational goals expressed by advisees, resulting in worse advice. When incentivized to be liked, advisers provide even more flattering advice, and advisees reward this by rating these advisers as more likable, despite the advice being less honest and less accurate (Study 2). The desire to avoid disappointment may lead to inequities if there are differences in expectations across social groups. In Study 3, we examine a setting in which men expect to perform better than women. We show that advisers take into account these expectations, leading to systematically different advice for men and women even when their gender is unknown to advisers. Advisers' efforts to avoid disappointment may thus contribute to systematic gender disparities in advice, with implications for downstream decision-making.

keywords          : "Gender difference, Advice, Interpresonal relationship"
wordcount         : "6,104"
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa7"
csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
classoption       : "man"
header-includes:
   - \usepackage{setspace, appendix, placeins, booktabs, tikz, siunitx}
   - \usepackage{tabularx, epigraph, float, colortbl, tabu, pdflscape}
   - \usepackage{tabularray}
   - \usetikzlibrary{positioning}
   - \raggedbottom
   - \usepackage[maxfloats=256]{morefloats}
   - \maxdeadcycles=1000
bibliography      : bibliography.bib
output: 
  # papaja::apa6_docx
  papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
always_allow_html: true
# bibliography: references.bib
---

```{r setup, include=FALSE}

@@

We present the results of three preregistered experiments in which
participants are paired anonymously as advisers and advisees. Advisers
are incentivized to offer accurate advice, with bonus earnings depending
on the outcome of an advisee's decision. We study a setting in which
advisees can communicate their expectations to an adviser, and
experimentally manipulate whether advisers observe this expectation. Our
design simulates organizational contexts where mentors and managers
often have some sense of an employee's confidence or aspirations and
thus could adjust their feedback to avoid disappointing them.
Interpersonal considerations would likely be stronger in situations in
which the adviser and advisee have an existing relationship and when the
advice is delivered face-to-face. Moreover, advisers usually do not
suffer any direct costs when their flattering advice leads to a poor
outcome for the recipient, particularly when it is not clear what would
have happened under a counterfactual.\^[In some cases, however, repeated
interaction might also increase incentives for honest feedback. Someone
who is known to persistently give overoptimistic advice may be viewed as
less trustworthy in the long run.]

@@

```{r study1regs}

@@

modelsummary::modelsummary(
  list(`(1)` = lm(Advice ~ ExpShow * Treatment, data = study1_advice), 
       `(2)` = lm(Advice ~ Treatment + Score, data = study1_advice %>% dplyr::filter(ExpShow == 1))),
  vcov = ~GiverID,
  coef_map = c(
    "TreatmentHigh" = "High Default",
    "ExpShow" = "Expectation Shown",
    "ExpShow:TreatmentHigh" = "Expectation x High Default",
    "Score" = "Performance",
    "(Intercept)" = "Constant"
  ),
  stars = TRUE,
  gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
  caption = "Advice to compete against the High Performer Group in Study 1. Displaying the non-binding choice of those for whom the Low Performer Group was selected by default makes it more likely that they are advised to compete against the High Performer Group (Column 1). Column 2 restricts the analysis to advisers who observed advisees' initial non-binding choice and controls for advisees' score on the quiz. Standard errors are clustered at the adviser level."
)

@@

We now examine whether advisers took the advisees' expectations into
account. Recall that advisees who were defaulted to competing against
the High Performer Group were indeed more likely to choose this group,
which, from the advisers' perspective, could signal high confidence in
their abilities. Our theory predicts that showing expectations should
lead to more flattering advice--- specifically, a recommendation to
compete against the High Performer Group--- for advisees with a high
default choice compared to those with a low default choice. Column 1 of
Table \@ref(tab:study1regs) reports a linear probability model with
advice to compete against the high group as the outcome measure,
incorporating the default choice advisees saw, whether expectations were
shown to the adviser, and the interaction of the two. Because each
adviser made ten recommendations, we cluster standard errors at the
adviser level. As predicted, we find a marginally significant
interaction effect: advisees with a default choice to compete against
the High Performer Group (nudged to have high expectations, though not
necessarily holding them) were more likely to receive advice to compete
against stronger competitors when their expectations were shown compared
to when they were hidden. We reported a linear probability model on
advising to compete against the High Performer group for advisers in the
"Expectations" treatment, controlling for the true performance of the
advisee in Column 2 of Table \@ref(tab:study1regs). We observe a
significant main effect of the default treatment, where a high default
increased the likelihood of receiving flattering advice. It is worth
noting that, since only
`r round(mean(study1_time1[study1_time1$default == "High",]$Guess),2)*100`
% of advisees in the high default group actually stuck with their
default choice, this represents a conservative test of the effect of
high expectations on the likelihood of receiving flattering advice.

```{r study1Advice, fig.cap ="Advisees who were nudged to express high expectations were more likely to receive advice to compete against the High Performer Group."}
study1_time2_advice %>%
  mutate(Treatment=factor(Treatment, levels=c('High','Low'), labels = c('High default', 'Low default'))) %>%
  group_by(Condition, Treatment) %>%
  summarize(avg_count = mean(Advice), ci = qnorm(0.975)*sd(Advice)/sqrt(dplyr::n())) %>%
  ggplot2::ggplot(ggplot2::aes(x = Condition, y = avg_count, fill = Treatment)) +
  ggplot2::geom_col(position = 'dodge') +
  ggplot2::geom_errorbar(ggplot2::aes(ymin = avg_count - ci, ymax = avg_count + ci), 
                         position = ggplot2::position_dodge(0.9), linewidth = 0.2, width = 0.1) +
  ggtitle("Advice: Meeting Expectations vs. Absent Expectations") +
  scale_fill_manual(values = c("#2A629A",'#FF7F3E')) +
  theme_bw() +
  labs(y = "P(Compete Against High Performer)", x = NULL) +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "grey"),
        axis.text.x = element_text(size = 12),
        legend.position = "none", # No legend
        plot.title = element_text(hjust = 0.5, size = 16)) +
  scale_x_discrete(labels = c("Expectation shown", "Expectation Absent")) +
   geom_text(aes(label=Treatment, y = -0.01), position = position_dodge(0.90))

@@

```{r}

@@

```{r}
library(dplyr)

@@

```{r study2advice, fig.cap="Advisers were asked to suggest what rank to bet on for someone they ranked as the 7th most attractive out of ten people. When incentivized for accuracy, most of them advised betting on 7, which was based on their evaluation of the recipient's attractiveness. However, when incentivized for likability, many participants gave flattering advice by suggesting betting on higher ranks that imply a higher evaluation of attractiveness."}

@@

Next, we turn our attention to the advisers (Figure
\@ref(fig:study2advice)). In the Accuracy condition, those uploading
selfies were advised to bet on rank
`r round(mean(study2_time2[study2_time2$Condition == "Accurate",]$Advice),2)`.
Notably, this is significantly more attractive than the 7th rank those
advisers had themselves guessed just on the prior screen
(`r papaja::apa_print(t.test(study2_time2[study2_time2$Condition == "Accurate",]$Advice, mu = 7, var.equal = T))$statistic`).
This suggests that even when incentivized for accuracy, participants
offered flattering advice.[^1] Importantly, and as predicted, we find
that advisers in the Likeable treatment recommend betting on a lower
rank, communicating that they think the participant in the selfie is
more attractive
(`r round(mean(study2_time2[study2_time2$Condition == "Likable",]$Advice),2)`,
`r papaja::apa_print(t.test(Advice ~ Condition, data = study2_time2, var.equal = T))$statistic`).
This suggests that advisers inferred that advising someone that they
were more attractive would make the adviser appear more likeable and
therefore provided advice that communicated a more favorable impression
of the participants' attractiveness. The distribution shown in Figure
\@ref(fig:study2advice) shows that participants do not simply tell
participants that they are the most attractive person in the group. They
may infer that flattering advise needs to be somewhat realistic to be
believable. We return to this in the general discussion.

@@

```{r study2regs}

modelsummary::modelsummary(list(`(1)` = lm(likable ~ Advice, data = study2_time3),
                                `(2)` = lm(warmAvg ~ Advice, data = study2_time3),
                                `(3)` = lm(trustAvg ~ Advice, data = study2_time3)),
                           vcov = c("classical", "classical", "classical"),
                           coef_map = c("Advice" = "Advised Rank",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
                           caption = "When individuals receive advice that implies a high level of attractiveness (lower rank), they tend to perceive the advice giver as more likable (Column 1), warm (Column 2). Column 3 shows that advisors are rated as more trustworthy when they advice lower ranks, but this relationship is only directional.")

@@

```{r study2rating, fig.cap="Participants tend to rate advisors who suggest betting on a higher rank (implying greater attractiveness) as more likable and warm."}

@@

```{r study2ratingByGroup, fig.cap="After participants received advice that suggested betting on a rank that was lower than, equal to, or higher than their self-expectations, they evaluated advisors based on their likability and warmth. Participants rated advisors who gave flattering advice and delivered a high evaluation of their attractiveness as more likable and warm than those whose advice implied a lower evaluation of their attractiveness."}

@@

As seen in Figure \@ref(fig:study2rating), we found that advisees who
suggested that the advisee was more attractive (lower rank) were indeed
rated as more likeable and warm (b = -0.109, p \< .05; b = -0.133, p \<
.001, respectively; See Columns 1 and 2 of Table \@ref(tab:study2regs)).
When using their self-expectations as a reference point, advisors who
suggest a rank lower than that point are perceived as less likable and
warm (see Figure \@ref(fig:study2ratingByGroup)). Interestingly, these
benefits are not at the cost of being hypocritical; advisers who
recommend a more favorable rank are viewed as no less trustworthy
(Column 3 of Table \@ref(tab:study2regs)).

@@

```{r study3regs}
study3_advice <- study3_advice %>% dplyr::mutate(Treatment = factor(Treatment, levels = c('Low', 'High'))) %>%
  dplyr::mutate(ExpShow = ifelse(Condition == 'E',1,0))

modelsummary::modelsummary(list(`(1)` = lm(Advice ~ Treatment + Performance , data = study3_advice %>% dplyr::filter(ExpShow == 1)),
                                `(2)` = lm(Advice ~ ExpShow*Gender , data = study3_advice),
                                `(3)` = lm(ExpectedBonus ~ ExpShow*Gender, data = study3_advice)),
                           vcov = c(~GiverID, ~GiverID,~GiverID),
                            coef_map = c("TreatmentHigh" = "High Expectation",
                                         "Performance" = "Performance",
                                         "ExpShow" = "Expectation Shown",
                                        "GenderMale" = "Advisee Male",
                                        "ExpShow:GenderMale" = "Expectation x Male",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
                           caption = "Column 1 displays the advice given to compete against a High Performance group based on whether the advisees were primed with low or high expectations and the expectation level shown to the advisor. Column 2 displays the advice given to compete against a High Performance group based on the gender of the targets and the expectation level shown to the advisor. Column 3 displays the expected bonus of the advice received based on the gender of the targets and the expectation level shown to the advisor. All standard errors clustered at the level of the advisor.")

@@

```{r study3genderdiff, fig.cap ="Advisers observed real performance and the participants' estimated performance, but not their gender. As a result of their higher expectations, men (blue) were advised to compete against the high performance group more often than women (red). The grey line shows advice absent expectations, which did not differ by gender."}
library(dplyr)
library(ggplot2)

@@

Next, we examine whether advisers took the advisee's expectations into
account. Column 1 of Table \@ref(tab:study3regs) reports a linear
probability model on advising to compete against the High Performer
group for advisers in the "Expectations" treatment, controlling for the
true performance of the advisee. Because each adviser made ten
recommendations, we cluster standard errors at the adviser level.
However, contrary to our expectations, we do not see a significant
effect of the expectations treatment. This may be because the induced
difference in expectations was too small.

However, recall that men were more confident in their performance than
were women. Our theory thus predicts that showing expectations should
lead to more flattering advice (i.e., a recommendation to compete
against the High Performer group) for men than for women. We report a
linear probability model with advice to compete against the high group
as the outcome measure, and the advisee's gender, whether expectations
were shown to the adviser, and the interaction of the two in Column 2 of
Table \@ref(tab:study3regs). As predicted, we find a significant
interaction effect: men are more likely to be advised to compete against
the High Performers when expectations are shown than when they are
hidden ($p < 0.001$). We show this result graphically in Figure
\@ref(fig:study3genderdiff)).

To examine the quality of advice, we computed the expected bonus
earnings for someone who followed the recommendations. For example, if
an adviser suggested competing against the High Performer group, we
matched the advisee against all 20 members of that group and determined
how often their score matched or exceeded that of the member. We then
multiplied this number by the respective bonus earnings (50 cents and 30
cents for the High and Low Performer group, respectively). To see if
including expectations leads to worse advise for men, we report a linear
probability model with the experimental treatment of the adviser, the
gender of the advisee, and their interaction in Column 3 of Table
\@ref(tab:study3regs). Displaying expectations led men to be advised to
compete against the High Performer group more often, and this advise
turned out to be bad: men receive worse advise than do women when
expectations are displayed, but not in their absence.[^3]

@@

```{r study3time3}

@@

modelsummary::modelsummary(list(`(1)` = lm(bonus ~ AdviseeCondition + AdviseeScore, data = study3_time3 %>%
                                             dplyr::filter(GiverCondition == 1)),
                                `(2)` = lm(bonus ~ Gender + AdviseeScore, data = study3_time3 %>%
                                             dplyr::filter(GiverCondition == 1)),
                                `(3)` = lm(Adopt ~ Gender + GiverCondition + Advice, data = study3_time3),
                                `(4)` = lm(Adopt ~ Gender + GiverCondition + Advice + Gender*GiverCondition + GiverCondition*Advice + Gender*Advice, data = study3_time3)),
                           vcov = c("classical", "classical","classical","classical"),
                            coef_map = c("AdviseeScore" = "Performance",
                                         "AdviseeConditionHigh" = "High Expectation",
                                        "GenderMale" = "Advisee Male",
                                        "GenderMale:GiverCondition" = "Expectation x Male",
                                        "GiverCondition" = "Expectation Shown",
                                        "Advice1" = "Advice: High Performer",
                                        "GenderMale:Advice1" = "Advice: High Performer x Male",
                                        "GiverCondition:Advice1" = "Expectation x Advice: High Performer",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
                           caption = "Column 1 displays the actual bonus based on the performance and whether the participant is primed with high expectations. Column 2 displays the actual bonus based on the performance and gender of the participant. Columns 3 and 4 display the chance of adopting the advice based on the gender of the participants, whether the advisee sees expectations, and whether the advice is to compete against high performers. The former considers only the main effect, while the latter also includes the interactive effect.")

@@

To determine whether flattering advice is truly costly, however, we need
to examine the outcome of the advisees. In particular, they could ignore
flattering advice, recognizing it as such and thus failing to adhere to
it. Align with our prediction, participants who were primed with high
expectations earned less when their expectations were conveyed to the
advisers, although this result is only directional (see Column 1 of
Table \@ref(tab:study3time3)). Similarly, as shown in Column 2 of Table
\@ref(tab:study3time3), we also found that male participants received
less reward when their expectations were presented, even when their
performance was the same. These findings suggest that flattering advice
is not without consequences. Furthermore, our research indicates that
because males are more likely to follow such advice (see Column 4 of
Table \@ref(tab:study3time3), their tendency towards competition
intensifies the cost of such advice.

@@

\begingroup
\singlespacing
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}

\noindent

@@

\endgroup

>>>>

---
title: "Flattering Advice: Avoiding Disappointment in Advice-Giving"
shorttitle: "Flattering Advice"
date: yes
author: 
  - name: "Amanda Chen"
    corresponding: true
    orcid: 0000-0002-4428-627X
    email: zchengj@connect.ust.hk
    affiliations:
      - id: id1
        name: "The Hong Kong University of Science and Technology"
        department: "Department of Management"
        city: "Hong Kong"
  - name: "David Hagmann"
    orcid: 0000-0002-2080-997X
    affiliations:
    - id: id1
author-note:
  disclosures:
    gratitude: The authors are grateful to XXX for their helpful comments.
abstract: Good advice improves decision quality but often requires delivering unpleasant truths that may disappoint advisees. Across three pre-registered and incentivized experiments involving real adviser-advisee interactions (N = 3,900), we show that advisers prioritize avoiding disappointment at the expense of accuracy and their own earnings. In Study 1, advisers financially rewarded for accuracy still tailor recommendations to aspirational goals expressed by advisees, resulting in worse advice. When incentivized to be liked, advisers provide even more flattering advice, and advisees reward this by rating these advisers as more likable, despite the advice being less honest and less accurate (Study 2). The desire to avoid disappointment may lead to inequities if there are differences in expectations across social groups. In Study 3, we examine a setting in which men expect to perform better than women. We show that advisers take into account these expectations, leading to systematically different advice for men and women even when their gender is unknown to advisers. Advisers' efforts to avoid disappointment may thus contribute to systematic gender disparities in advice, with implications for downstream decision-making.

keywords: [Gender difference, Advice, Interpresonal relationship]
bibliography: bibliography.bib
floatsintext: true
mask: false
link-citations: false
format:
  apaquarto-html: default
  apaquarto-pdf:
    documentmode: man
    keep-tex: true
    prefer-html: true
  apaquarto-docx:
    default: true
    prefer-html: true
---

```{r}
#| label: setup
#| include: false

@@

We present the results of three preregistered experiments in which participants are paired anonymously as advisers and advisees. Advisers are incentivized to offer accurate advice, with bonus earnings depending on the outcome of an advisee's decision. We study a setting in which advisees can communicate their expectations to an adviser, and experimentally manipulate whether advisers observe this expectation. Our design simulates organizational contexts where mentors and managers often have some sense of an employee's confidence or aspirations and thus could adjust their feedback to avoid disappointing them. Interpersonal considerations would likely be stronger in situations in which the adviser and advisee have an existing relationship and when the advice is delivered face-to-face. Moreover, advisers usually do not suffer any direct costs when their flattering advice leads to a poor outcome for the recipient, particularly when it is not clear what would have happened under a counterfactual.\^\[In some cases, however, repeated interaction might also increase incentives for honest feedback. Someone who is known to persistently give overoptimistic advice may be viewed as less trustworthy in the long run.\]

@@

```{r}
#| label: tbl-study1regs
#| tbl-cap: "Advice to compete against the High Performer Group in Study 1. Displaying the non-binding choice of those for whom the Low Performer Group was selected by default makes it more likely that they are advised to compete against the High Performer Group (Column 1). Column 2 restricts the analysis to advisers who observed advisees' initial non-binding choice and controls for advisees' score on the quiz. Standard errors are clustered at the adviser level."

@@

modelsummary::modelsummary(
  list(`(1)` = lm(Advice ~ ExpShow * Treatment, data = study1_advice), 
       `(2)` = lm(Advice ~ Treatment + Score, data = study1_advice %>% dplyr::filter(ExpShow == 1))),
  vcov = ~GiverID,
  coef_map = c(
    "TreatmentHigh" = "High Default",
    "ExpShow" = "Expectation Shown",
    "ExpShow:TreatmentHigh" = "Expectation x High Default",
    "Score" = "Performance",
    "(Intercept)" = "Constant"
  ),
  stars = TRUE,
  gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0))
)

@@

We now examine whether advisers took the advisees' expectations into account. Recall that advisees who were defaulted to competing against the High Performer Group were indeed more likely to choose this group, which, from the advisers' perspective, could signal high confidence in their abilities. Our theory predicts that showing expectations should lead to more flattering advice--- specifically, a recommendation to compete against the High Performer Group--- for advisees with a high default choice compared to those with a low default choice. Column 1 of @tbl-study1regs reports a linear probability model with advice to compete against the high group as the outcome measure, incorporating the default choice advisees saw, whether expectations were shown to the adviser, and the interaction of the two. Because each adviser made ten recommendations, we cluster standard errors at the adviser level. As predicted, we find a marginally significant interaction effect: advisees with a default choice to compete against the High Performer Group (nudged to have high expectations, though not necessarily holding them) were more likely to receive advice to compete against stronger competitors when their expectations were shown compared to when they were hidden. We reported a linear probability model on advising to compete against the High Performer group for advisers in the "Expectations" treatment, controlling for the true performance of the advisee in Column 2 of @tbl-study1regs. We observe a significant main effect of the default treatment, where a high default increased the likelihood of receiving flattering advice. It is worth noting that, since only `r round(mean(study1_time1[study1_time1$default == "High",]$Guess),2)*100` % of advisees in the high default group actually stuck with their default choice, this represents a conservative test of the effect of high expectations on the likelihood of receiving flattering advice.

```{r}
#| label: fig-study1Advice
#| fig-cap: "Advisees who were nudged to express high expectations were more likely to receive advice to compete against the High Performer Group."

study1_time2_advice %>%
  mutate(Treatment=factor(Treatment, levels=c('High','Low'), labels = c('High default', 'Low default'))) %>%
  group_by(Condition, Treatment) %>%
  summarize(avg_count = mean(Advice), ci = qnorm(0.975)*sd(Advice)/sqrt(dplyr::n())) %>%
  ggplot2::ggplot(ggplot2::aes(x = Condition, y = avg_count, fill = Treatment)) +
  ggplot2::geom_col(position = 'dodge') +
  ggplot2::geom_errorbar(ggplot2::aes(ymin = avg_count - ci, ymax = avg_count + ci), 
                         position = ggplot2::position_dodge(0.9), linewidth = 0.2, width = 0.1) +
  ggtitle("Advice: Meeting Expectations vs. Absent Expectations") +
  scale_fill_manual(values = c("#2A629A",'#FF7F3E')) +
  theme_bw() +
  labs(y = "P(Compete Against High Performer)", x = NULL) +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "grey"),
        axis.text.x = element_text(size = 12),
        legend.position = "none", # No legend
        plot.title = element_text(hjust = 0.5, size = 16)) +
  scale_x_discrete(labels = c("Expectation shown", "Expectation Absent")) +
   geom_text(aes(label=Treatment, y = -0.01), position = position_dodge(0.90))

@@

```{r}
#| include: false

@@

```{r}
#| include: false
library(dplyr)

@@

```{r}
#| label: fig-study2advice
#| fig-cap: "Advisers were asked to suggest what rank to bet on for someone they ranked as the 7th most attractive out of ten people. When incentivized for accuracy, most of them advised betting on 7, which was based on their evaluation of the recipient's attractiveness. However, when incentivized for likability, many participants gave flattering advice by suggesting betting on higher ranks that imply a higher evaluation of attractiveness."

@@

Next, we turn our attention to the advisers (@fig-study2advice). In the Accuracy condition, those uploading selfies were advised to bet on rank `r round(mean(study2_time2[study2_time2$Condition == "Accurate",]$Advice),2)`. Notably, this is significantly more attractive than the 7th rank those advisers had themselves guessed just on the prior screen (`r papaja::apa_print(t.test(study2_time2[study2_time2$Condition == "Accurate",]$Advice, mu = 7, var.equal = T))$statistic`). This suggests that even when incentivized for accuracy, participants offered flattering advice.[^1] Importantly, and as predicted, we find that advisers in the Likeable treatment recommend betting on a lower rank, communicating that they think the participant in the selfie is more attractive (`r round(mean(study2_time2[study2_time2$Condition == "Likable",]$Advice),2)`, `r papaja::apa_print(t.test(Advice ~ Condition, data = study2_time2, var.equal = T))$statistic`). This suggests that advisers inferred that advising someone that they were more attractive would make the adviser appear more likeable and therefore provided advice that communicated a more favorable impression of the participants' attractiveness. The distribution shown in @fig-study2advice shows that participants do not simply tell participants that they are the most attractive person in the group. They may infer that flattering advise needs to be somewhat realistic to be believable. We return to this in the general discussion.

@@

```{r}
#| label: tbl-study2regs
#| tbl-cap: "When individuals receive advice that implies a high level of attractiveness (lower rank), they tend to perceive the advice giver as more likable (Column 1), warm (Column 2). Column 3 shows that advisors are rated as more trustworthy when they advice lower ranks, but this relationship is only directional."

modelsummary::modelsummary(list(`(1)` = lm(likable ~ Advice, data = study2_time3),
                                `(2)` = lm(warmAvg ~ Advice, data = study2_time3),
                                `(3)` = lm(trustAvg ~ Advice, data = study2_time3)),
                           vcov = c("classical", "classical", "classical"),
                           coef_map = c("Advice" = "Advised Rank",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)))

@@

```{r}
#| label: fig-study2rating
#| fig-cap: "Participants tend to rate advisors who suggest betting on a higher rank (implying greater attractiveness) as more likable and warm."

@@

```{r}
#| label: fig-study2ratingByGroup
#| fig-cap: "After participants received advice that suggested betting on a rank that was lower than, equal to, or higher than their self-expectations, they evaluated advisors based on their likability and warmth. Participants rated advisors who gave flattering advice and delivered a high evaluation of their attractiveness as more likable and warm than those whose advice implied a lower evaluation of their attractiveness."

@@

As seen in @fig-study2rating, we found that advisees who suggested that the advisee was more attractive (lower rank) were indeed rated as more likeable and warm (b = -0.109, p \< .05; b = -0.133, p \< .001, respectively; See Columns 1 and 2 of @tbl-study2regs). When using their self-expectations as a reference point, advisors who suggest a rank lower than that point are perceived as less likable and warm (see @fig-study2ratingByGroup). Interestingly, these benefits are not at the cost of being hypocritical; advisers who recommend a more favorable rank are viewed as no less trustworthy (Column 3 of @tbl-study2regs).

@@

```{r}
#| label: tbl-study3regs
#| tbl-cap: "Column 1 displays the advice given to compete against a High Performance group based on whether the advisees were primed with low or high expectations and the expectation level shown to the advisor. Column 2 displays the advice given to compete against a High Performance group based on the gender of the targets and the expectation level shown to the advisor. Column 3 displays the expected bonus of the advice received based on the gender of the targets and the expectation level shown to the advisor. All standard errors clustered at the level of the advisor."

study3_advice <- study3_advice %>% dplyr::mutate(Treatment = factor(Treatment, levels = c('Low', 'High'))) %>%
  dplyr::mutate(ExpShow = ifelse(Condition == 'E',1,0))

modelsummary::modelsummary(list(`(1)` = lm(Advice ~ Treatment + Performance , data = study3_advice %>% dplyr::filter(ExpShow == 1)),
                                `(2)` = lm(Advice ~ ExpShow*Gender , data = study3_advice),
                                `(3)` = lm(ExpectedBonus ~ ExpShow*Gender, data = study3_advice)),
                           vcov = c(~GiverID, ~GiverID,~GiverID),
                            coef_map = c("TreatmentHigh" = "High Expectation",
                                         "Performance" = "Performance",
                                         "ExpShow" = "Expectation Shown",
                                        "GenderMale" = "Advisee Male",
                                        "ExpShow:GenderMale" = "Expectation x Male",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)))

@@

```{r}
#| label: fig-study3genderdiff
#| fig-cap: "Advisers observed real performance and the participants' estimated performance, but not their gender. As a result of their higher expectations, men (blue) were advised to compete against the high performance group more often than women (red). The grey line shows advice absent expectations, which did not differ by gender."

library(dplyr)
library(ggplot2)

@@

Next, we examine whether advisers took the advisee's expectations into account. Column 1 of @tbl-study3regs reports a linear probability model on advising to compete against the High Performer group for advisers in the "Expectations" treatment, controlling for the true performance of the advisee. Because each adviser made ten recommendations, we cluster standard errors at the adviser level. However, contrary to our expectations, we do not see a significant effect of the expectations treatment. This may be because the induced difference in expectations was too small.

However, recall that men were more confident in their performance than were women. Our theory thus predicts that showing expectations should lead to more flattering advice (i.e., a recommendation to compete against the High Performer group) for men than for women. We report a linear probability model with advice to compete against the high group as the outcome measure, and the advisee's gender, whether expectations were shown to the adviser, and the interaction of the two in Column 2 of @tbl-study3regs. As predicted, we find a significant interaction effect: men are more likely to be advised to compete against the High Performers when expectations are shown than when they are hidden ($p < 0.001$). We show this result graphically in @fig-study3genderdiff.

To examine the quality of advice, we computed the expected bonus earnings for someone who followed the recommendations. For example, if an adviser suggested competing against the High Performer group, we matched the advisee against all 20 members of that group and determined how often their score matched or exceeded that of the member. We then multiplied this number by the respective bonus earnings (50 cents and 30 cents for the High and Low Performer group, respectively). To see if including expectations leads to worse advise for men, we report a linear probability model with the experimental treatment of the adviser, the gender of the advisee, and their interaction in Column 3 of @tbl-study3regs. Displaying expectations led men to be advised to compete against the High Performer group more often, and this advise turned out to be bad: men receive worse advise than do women when expectations are displayed, but not in their absence.[^3]

@@

```{r}
#| label: tbl-study3time3
#| tbl-cap: "Column 1 displays the actual bonus based on the performance and whether the participant is primed with high expectations. Column 2 displays the actual bonus based on the performance and gender of the participant. Columns 3 and 4 display the chance of adopting the advice based on the gender of the participants, whether the advisee sees expectations, and whether the advice is to compete against high performers. The former considers only the main effect, while the latter also includes the interactive effect."

@@

modelsummary::modelsummary(list(`(1)` = lm(bonus ~ AdviseeCondition + AdviseeScore, data = study3_time3 %>%
                                             dplyr::filter(GiverCondition == 1)),
                                `(2)` = lm(bonus ~ Gender + AdviseeScore, data = study3_time3 %>%
                                             dplyr::filter(GiverCondition == 1)),
                                `(3)` = lm(Adopt ~ Gender + GiverCondition + Advice, data = study3_time3),
                                `(4)` = lm(Adopt ~ Gender + GiverCondition + Advice + Gender*GiverCondition + GiverCondition*Advice + Gender*Advice, data = study3_time3)),
                           vcov = c("classical", "classical","classical","classical"),
                            coef_map = c("AdviseeScore" = "Performance",
                                         "AdviseeConditionHigh" = "High Expectation",
                                        "GenderMale" = "Advisee Male",
                                        "GenderMale:GiverCondition" = "Expectation x Male",
                                        "GiverCondition" = "Expectation Shown",
                                        "Advice1" = "Advice: High Performer",
                                        "GenderMale:Advice1" = "Advice: High Performer x Male",
                                        "GiverCondition:Advice1" = "Expectation x Advice: High Performer",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)))

@@

To determine whether flattering advice is truly costly, however, we need to examine the outcome of the advisees. In particular, they could ignore flattering advice, recognizing it as such and thus failing to adhere to it. Align with our prediction, participants who were primed with high expectations earned less when their expectations were conveyed to the advisers, although this result is only directional (see Column 1 of @tbl-study3time3). Similarly, as shown in Column 2 of @tbl-study3time3, we also found that male participants received less reward when their expectations were presented, even when their performance was the same. These findings suggest that flattering advice is not without consequences. Furthermore, our research indicates that because males are more likely to follow such advice (see Column 4 of @tbl-study3time3), their tendency towards competition intensifies the cost of such advice.

@@

@@
